---
title: "race political ideology"
author: "Yiyang Shu"
date: "2026-10-07"
output:
  pdf_document:
    latex_engine: xelatex
header-includes:
  - \usepackage{fvextra}
  - \RecustomVerbatimEnvironment{Highlighting}{Verbatim}{breaklines=true,breakanywhere=true,commandchars=\\\{\}}
---

Extract the data for distribution analysis
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)
library(tidyverse)

anes_full <- read_csv("ANES_Panel_Full.csv")
```

```{r}
anes_distribution <- anes_full %>%
  select(
    # Respondent ID
    V160001,
    V160001_orig,

    #PRE Party of Registration
    V161019,   # 2016
    V201018,   # 2020
    V241025,   # 2024

    #POST Party of Registration
    V162030,   # 2016
    V202064,   # 2020
    V242064,   # 2024

    #Race / Ethnicity
    V161310x,  # 2016
    V201549x,  # 2020
    V241501x,  # 2024

    #Liberal-Conservative Self-Placement PRE
    V161126,   # 2016
    V201200,   # 2020
    V241177,   # 2024

    #2016 POST ideology
    V162171
  )
```

So the codes we have:
PRE Party Registration
2016 - V161019
2020 - V201018
2024 - V241025

PRE Race/Ethnicity
2016 - V161310x
2020 - V201549x
2024 - V241501x

PRE Political Ideology / Liberal-Conservative Self-Placement
2016 - V161126
2020 - V201200
2024 - V241177

POST Party Registration
2016 - V162030
2020 - V202064
2024 - V242064

POST Political Ideology / Liberal-Conservative Self-Placement
2016 - V162171
2020 - No corresponding repeated POST measure identified
2024 - No corresponding repeated POST measure identified

```{r}
#write_csv(
#  anes_distribution,
#  "anes_distribution_race_party_liberal.csv"
#)
```

```{r}
#Set Common plot theme
plot_theme <- theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    plot.title = element_text(face = "bold")
  )
```

```{r}
#1.Pre Party of Registration
pre_party_dist <- anes_distribution %>%
  select(
    `2016` = V161019,
    `2020` = V201018,
    `2024` = V241025
  ) %>%
  pivot_longer(
    everything(),
    names_to = "year",
    values_to = "party"
  ) %>%
  filter(party %in% c(1, 2, 4, 5)) %>%
  mutate(
    year = factor(year, levels = c("2016", "2020", "2024")),
    party = factor(
      party,
      levels = c(1, 2, 4, 5),
      labels = c(
        "Democratic Party",
        "Republican Party",
        "None / Independent",
        "Other Party"
      )
    )
  ) %>%
  count(year, party, name = "n") %>%
  group_by(year) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ungroup()

pre_party_dist
```

```{r}
#Draw the plot
ggplot(
  pre_party_dist,
  aes(x = party, y = percent, fill = year)
) +
  geom_col(position = "dodge") +
  labs(
    title = "PRE Party of Registration Distribution by Year",
    x = "Party of Registration",
    y = "Percent",
    fill = "Year"
  ) +
  plot_theme
```

```{r}
#2.Race / Ethnicity 
race_long <- anes_distribution %>%
  select(
    `2016` = V161310x,
    `2020` = V201549x,
    `2024` = V241501x
  ) %>%
  pivot_longer(
    everything(),
    names_to = "year",
    values_to = "race_raw"
  ) %>%
  mutate(
    race = case_when(

      # 2016 coding
      year == "2016" & race_raw == 1 ~ "White, non-Hispanic",
      year == "2016" & race_raw == 2 ~ "Black, non-Hispanic",
      year == "2016" & race_raw == 5 ~ "Hispanic",
      year == "2016" & race_raw == 3 ~ "Asian/Pacific Islander",
      year == "2016" & race_raw == 4 ~ "Native American/Other",
      year == "2016" & race_raw == 6 ~ "Multiple races",

      # 2020 and 2024 coding
      year %in% c("2020", "2024") & race_raw == 1 ~
        "White, non-Hispanic",

      year %in% c("2020", "2024") & race_raw == 2 ~
        "Black, non-Hispanic",

      year %in% c("2020", "2024") & race_raw == 3 ~
        "Hispanic",

      year %in% c("2020", "2024") & race_raw == 4 ~
        "Asian/Pacific Islander",

      year %in% c("2020", "2024") & race_raw == 5 ~
        "Native American/Other",

      year %in% c("2020", "2024") & race_raw == 6 ~
        "Multiple races",

      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(race)) %>%
  mutate(
    year = factor(
      year,
      levels = c("2016", "2020", "2024")
    ),
    race = factor(
      race,
      levels = c(
        "White, non-Hispanic",
        "Black, non-Hispanic",
        "Hispanic",
        "Asian/Pacific Islander",
        "Native American/Other",
        "Multiple races"
      )
    )
  )


race_dist <- race_long %>%
  count(year, race, name = "n") %>%
  group_by(year) %>%
  mutate(
    percent = n / sum(n) * 100
  ) %>%
  ungroup()

race_dist
```

```{r}
ggplot(
  race_dist,
  aes(x = race, y = percent, fill = year)
) +
  geom_col(position = "dodge") +
  labs(
    title = "PRE Race/Ethnicity Distribution by Year",
    x = "Race/Ethnicity",
    y = "Percent",
    fill = "Year"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```

```{r}
#3. PRE Liberal–Conservative Self-Placement
pre_ideology_dist <- anes_distribution %>%
  select(
    `2016` = V161126,
    `2020` = V201200,
    `2024` = V241177
  ) %>%
  pivot_longer(
    everything(),
    names_to = "year",
    values_to = "ideology"
  ) %>%
  filter(ideology %in% 1:7) %>%
  mutate(
    year = factor(year, levels = c("2016", "2020", "2024")),
    ideology = factor(
      ideology,
      levels = 1:7,
      labels = c(
        "Extremely liberal",
        "Liberal",
        "Slightly liberal",
        "Moderate",
        "Slightly conservative",
        "Conservative",
        "Extremely conservative"
      )
    )
  ) %>%
  count(year, ideology, name = "n") %>%
  group_by(year) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ungroup()

pre_ideology_dist
```

```{r}
ggplot(
  pre_ideology_dist,
  aes(x = ideology, y = percent, fill = year)
) +
  geom_col(position = "dodge") +
  labs(
    title = "PRE Political Ideology Distribution by Year",
    x = "Liberal-Conservative Self-Placement",
    y = "Percent",
    fill = "Year"
  ) +
  plot_theme
```

```{r}
#4.POST Party of Registration (this one has many responses missing in 2024)
post_party_dist <- anes_distribution %>%
  select(
    `2016` = V162030,
    `2020` = V202064,
    `2024` = V242064
  ) %>%
  pivot_longer(
    everything(),
    names_to = "year",
    values_to = "party"
  ) %>%
  filter(party %in% c(1, 2, 4, 5)) %>%
  mutate(
    year = factor(year, levels = c("2016", "2020", "2024")),
    party = factor(
      party,
      levels = c(1, 2, 4, 5),
      labels = c(
        "Democratic Party",
        "Republican Party",
        "None / Independent",
        "Other Party"
      )
    )
  ) %>%
  count(year, party, name = "n") %>%
  group_by(year) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ungroup()

post_party_dist
```

```{r}
ggplot(
  post_party_dist,
  aes(x = party, y = percent, fill = year)
) +
  geom_col(position = "dodge") +
  labs(
    title = "POST Party of Registration Distribution by Year",
    x = "Party of Registration",
    y = "Percent",
    fill = "Year"
  ) +
  plot_theme
```



