---
title: "687 Group 7 Project"
author: "Mi Huynh"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

**Codebook:**

**V200011a** - 2016-2020 panel pre-post-election weight V202144
(0-100) - Feeling thermometer for Republican candidate Trump

**V201200** - PRE: 7PT SCALE LIBERAL-CONSERVATIVE SELFPLACEMENT -9.
Refused -8. Don’t know 1. Extremely liberal 2. Liberal 3. Slightly
liberal 4. Moderate; middle of the road 5. Slightly conservative 6.
Conservative 7. Extremely conservative 99. Haven’t thought much about
this

**V202064** - POST: Election party of registration -9. Refused -8. Don’t
know -7. No post-election data, deleted due to incomplete interview -6.
No post-election interview -1. Inapplicable 1. Democratic party 2.
Republican party 4. None or ‘independent’ 5. Other party {SPECIFY}

**V202637** - POST: IWR OBS: RESPONDENT’S GENDER -8. Don’t know -7. No
post-election data, deleted due to incomplete interview -6. No
post-election interview -5. Interview breakoff (sufficient partial IW)
-1. Inapplicable 1. Male 2. Female

**V201549x** - PRE: SUMMARY: R SELF-IDENTIFIED RACE/ETHNICITY -9.
Refused -8. Don’t know 1. White, non-Hispanic 2. Black, non-Hispanic 3.
Hispanic 4. Asian or Native Hawaiian/other Pacific Islander,
non-Hispanic alone 5. Native American/Alaska Native or other race,
non-Hispanic alone 6. Multiple races, non-Hispanic

**V203000** - SAMPLE - FIPS codes 1. Alabama 2. Alaska 4. Arizona 5.
Arkansas 6. California 8. Colorado 9. Connecticut 10. Delaware 11.
Washington DC 12. Florida 13. Georgia 15. Hawaii 16. Idaho 17. Illinois
18. Indiana 19. Iowa 20. Kansas 21. Kentucky 22. Louisiana 23. Maine 24.
Maryland 25. Massachusetts 26. Michigan 27. Minnesota 28. Mississippi
29. Missouri 30. Montana 31. Nebraska 32. Nevada 33. New Hampshire 34.
New Jersey 35. New Mexico 36. New York 37. North Carolina 38. North
Dakota 39. Ohio 40. Oklahoma 41. Oregon 42. Pennsylvania 44. Rhode
Island 45. South Carolina 46. South Dakota 47. Tennessee 48. Texas 49.
Utah 50. Vermont 51. Virginia 53. Washington 54. West Virginia 55.
Wisconsin 56. Wyoming

```{r, warnings=FALSE}
install.packages("readr")
install.packages("dplyr")
library(readr)
library(dplyr)
```

```{r}
anes_mergedfile_2016_2020_2024 <- read_csv("~/Library/Mobile Documents/com~apple~CloudDocs/Michigan Program in Survey and Data Science/FA 2026/SURVMETH 687- Applications of Statistical Modeling/Survmeth 687- Group 7 Project/anes_mergedfile_2016-2020-2024panel_20260519 3.csv")
anes_cleaned <- anes_mergedfile_2016_2020_2024 %>% select(contains(c("V200011a", "V202144","V201200","V202064","V202637","V201549x","V203000")))
write_csv(anes_cleaned, file = "anes_cleaned_0925.csv")
```
